SunPy - Python for Solar Physics

### 2014-01-27 3-1-VSO.tex
Edited the VSO section

The Virtual Solar Observatory (VSO) was built to facilitate
access to solar data from the different archives around the 
world \cite{hill_virtual_2009}.  
Enabling therefore a single standard query interface to all
the available products from solar observation.  
VSO allows to query data not just by instrument or mission for 
a time range but by physical variables as for example:
intensity, magnetic field, particle flux or spectral range to name a few.
VSO's main interface is web-based\footnote{\url{http://virtualsolar.org}}, 
however a SOAP webservice is also available.
SunPy access to the VSO through the \texttt{vso} module which it uses 
the \texttt{suds} package to connect to the webservice.
This module was developed under the ESA's Summer of Code programme of 2010.

The \texttt{vso} module includes advanced features through the use
of VSO attributes objects.  
The time range, the instrument, the wavelength, the physical observable,
and many other properties are set as attributes objects which could
be combine using \texttt{and} - just by chaining together the arguments -
or \texttt{or} operators - by joining the attributes using the ``$|$'' operator.
Listing~\ref{code:vso_query} shows few examples of queries made using
such attributes.

\begin{listing}[h]
\begin{minted}{pycon}
>>> from sunpy.net import vso
>>> client=vso.VSOClient()
>>> tstart, tend = '2011/6/7 05:30', '2011/6/7 10:30'

Querying just a single instrument
>>> lasco_query = client.query(vso.attrs.Time(tstart, tend), 
...                            vso.attrs.Instrument('lasco'))
>>> lasco_query.num_records()
40
>>> lasco_query.show() 
Start time           End time             Source  Instrument  Type  
----------           --------             ------  ----------  ----  
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO    LASCO       CORONA
2011-06-07 05:43:09  2011-06-07 05:43:29  SOHO    LASCO       CORONA
...

Adding to the query STEREO's coronographs: COR1
>>> instrument_condition = ( vso.attrs.Instrument('lasco') | 
...                          vso.attrs.Detector('cor1') )
>>> coronographs = client.query(vso.attrs.Time(tstart, tend)), 
...                             instrument_condition)
>>> coronographs.num_records()
485
>>> coronographs.show() 
Start time           End time             Source    Instrument  Type  
----------           --------             ------    ----------  ----  
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO      LASCO       CORONA
2011-06-07 05:43:09  2011-06-07 05:43:29  SOHO      LASCO       CORONA
...
2011-06-07 10:25:43  2011-06-07 10:25:45  STEREO_B  SECCHI      CORONA
2011-06-07 10:30:00  2011-06-07 10:30:01  STEREO_A  SECCHI      CORONA
\end{minted}
\caption{Examples of different advanced queries through the \texttt{vso} 
  module using the attributes objects.}
\label{code:vso_query}
\end{listing}

The \texttt{vso} module also provides an interface to download the data,
allowing to customise the directory structure to save the data.
Listing~\ref{code:vso_down} exemplifies such case.

\begin{listing}[h]
\begin{minted}{pycon}
>>> pathformat = '/data/{instrument}/{detector}/{file}.fits'
>>> results = client.get(coronographs, 
...                      path = pathformat)

\end{minted}
\caption{From listing~\ref{code:vso_query}, this examples shows the 
  download method of the \texttt{vso} module.}
\label{code:vso_down}
\end{listing}

>>>>

The Virtual Solar Observatory (VSO) provides a single, standard query
interface to access solar data from different archives around the world
 \cite{hill2009}.
Data products can be requested for specific instruments or missions, but
can also be requested based on physical parameters of the data product such
as the wavelength range.
In addition to the VSO's primary, web-based interface
(\url{http://virtualsolar.org}), a SOAP (Simple Object Access Protocol)
service is also available.
SunPy's \texttt{vso} module accesses the VSO via the SOAP service using the
\texttt{suds} package.
This module was developed under the ESA's Summer of Code programme of 2011.

Listing~\ref{code:vso_query} shows a number of examples of how to make queries
using the \texttt{vso} module.
Each query is constructed by one or more attributes objects, where each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, wavelength, or  other properties.
Two attribute objects can be combined with a logical \texttt{or} operation
using the ``$|$'' operator.
All attribute objects provided to the query as keyword arguments are combined
with a logical \texttt{and} operation.

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import vso
>>> client = vso.VSOClient()
>>> tstart, tend = '2011/6/7 05:30', '2011/6/7 05:45'

Querying a single instrument over a time range
>>> lasco_query = client.query(vso.attrs.Time(tstart, tend),
...                            vso.attrs.Instrument('lasco'))
>>> lasco_query.num_records()
2
>>> lasco_query.show() 
Start time           End time             Source  Instrument  Type
----------           --------             ------  ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO    LASCO       CORONA
2011-06-07 05:43:09  2011-06-07 05:43:29  SOHO    LASCO       CORONA

Querying multiple instruments/missions over the same time range
>>> instrument_condition = ( vso.attrs.Instrument('lasco') |
...                          vso.attrs.Detector('cor1') )
>>> coronagraphs = client.query(vso.attrs.Time(tstart, tend),
...                             instrument_condition)
>>> coronagraphs.num_records()
30
>>> coronographs.show() 
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO      LASCO       CORONA
2011-06-07 05:43:09  2011-06-07 05:43:29  SOHO      LASCO       CORONA
2011-06-07 05:30:00  2011-06-07 05:30:01  STEREO_A  SECCHI      CORONA
...
2011-06-07 05:43:13  2011-06-07 05:43:15  STEREO_B  SECCHI      CORONA
2011-06-07 05:45:00  2011-06-07 05:45:01  STEREO_A  SECCHI      CORONA

Querying two wavelength ranges over the same time range
>>> wavelength_condition = ( vso.attrs.Wave(125, 135) |
...                          vso.attrs.Wave(165, 175) ) # in angstroms
>>> euv = client.query(vso.attrs.Time(tstart, tend),
...                    wavelength_condition)
>>> euv.num_records()
199
>>> euv.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 00:00:00  2011-06-08 00:00:00  SDO       EVE         FULLDISK
...
2011-06-07 05:31:09  2011-06-07 05:31:19  PROBA2    SWAP        FULLSUN
...
2011-06-07 05:45:00  2011-06-07 05:45:01  SDO       AIA         FULLDISK
\end{minted}
\caption{Examples of different VSO queries using attribute objects.}
\label{code:vso_query}
\end{listing}

Listing~\ref{code:vso_down} shows how to download the data using the result
from the query.
One can use the metadata associated with the data sets to specify how the
files will be named locally, such as creating subdirectories based on the
instrument names.

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
>>> pathformat = '/data/{instrument}/{detector}/{file}.fits'
>>> results = client.get(coronagraphs, path = pathformat)
\end{minted}
\caption{Example of how to download data using a query result (see
listing~\ref{code:vso_query}).}
\label{code:vso_down}
\end{listing}
